Quantum internet \& quantum machine learning

### 2019-03-11 Nana QML section/QinternetQML.tex
minor

\documentclass[twocolumn, aps, rmp, amsmath, amssymb, nofootinbib, superscriptaddress, longbibliography, floatfix, table-of-contents, eqsecnum]{revtex4-2}

@@

Today classical machine learning is affecting the understanding and better regulation of classical network systems which can be of use in the modern internet. These include pattern recognition for individual computing devices, security and fault management, routing and traffic management, resource management and distributed computation. In these applications, it is not only the speed of processing that is important, but reliability and security can be paramount.

The large intersection between machine learning and network systems (the internet included) is perhaps not surprising. Firstly, machine learning relies on access to data and in many real-life applications, data naturally arise from distributed sources. Secondly, especially for complex systems like large networks, the information to process is complex and contains many uncertainties and errors. Exactly solvable models in these regimes are few and far between, and machine learning may be helpful in making predictions in these messy environments.

In the coming quantum age we can envision three possible ways where quantum resources can come into our internet: quantum communication, quantum processing at individual nodes and data that is naturally quantum in origin. To begin, we first make a classification of the four different types of networks we can have: see Table~\eqref{tab:CommTable}. We can classify them into:

\begin{itemize}
\item CC: classical data and processing over classical network, or the current internet.
\item CQ: classical data and processing over quantum network.
\item QC: quantum data and processing over classical network.
\item QQ: quantum data and processing over quantum network, or a fully-quantum internet.
\end{itemize}

So now we can pose the inevitable question: how can these different quantum resources in a network influence the relationship between the internet and machine learning?

This is not yet an active research area in its own right. However, there are some preliminary toolkits we can consider that are starting to be developed in the new field of quantum machine learning. By first summarising the intersection between the classical internet and classical machine learning, it gives us an idea of the kinds of tools we need to begin examining their quantum counterparts. For instance, we will see how aspects of quantum processing of quantum data might be aided by machine learning, how machine learning may be enhanced by quantum resources and also in turn how machine learning may be implemented in these distributed quantum settings. See Table~\eqref{tab:CommTable2} for a brief overview.

%It is natural to consider machine learning when we have an internet. Firstly, in machine learning applications, data naturally arise from distributed sources, so we begin from a network setting. There are also four main requirements we have in network settings (i) security (ii) speed and reliability of device information processing at each node (iii) speed and reliability of routing protocols (iv) speed and reliability of data transmission. Some of these may benefit from machine learning techniques. We then introduce quantum resources for two main reasons. The first is when the data we want to process is naturally quantum, so we wish to use machine learning to either diminish our resources to process this data or want to perform machine learning algorithms directly on this data. The second reason is that we exploit quantum resources to enhance the processing of classical data by means of machine learning.
%There are 3 parts to machine learning: data collection (offline or online); feature engineering and model learning.
\section{Classical machine learning in classical networks}
 To create and maintain an efficient classical network, one requires efficient and reliable processing on individual nodes, secure processing, efficient routing and data transmission, efficient uses of resources and a means for distributed information processing. For overview of methods see for example \cite{boutaba2018comprehensive}, \cite{wang2018machine}.  We give a brief overview on how machine learning can be used in each of these areas.

\subsection{Machine learning basics for individual processors}
Machine learning algorithms enable one to make predictions about one's current or future dataset without requiring explicit instructions. Since its aims are directed more towards \textit{prediction} rather than purely \textit{estimation}, it differs from the field of statistical estimation even though it shares many common tools.

There are 3 mains paradigms for machine learning: supervised, unsupervised and reinforcement learning. Supervised learning relies on having training data from which to form inferences and to make predictions about new incoming `test' data. On the other hand, unsupervised learning algorithms make inferences using the data at hand without access to a training stage, much like a student learning without an instructor. Reinforcement learning operates using a different framework and aims to find the best action steps in a particular environment to maximise a given reward.

Machine learning is used regularly for data collection, feature engineering and model learning. There are many excellent references on this broad subject, for instance see the introductory texts \cite{bishop2006pattern, shalev2014understanding, trevor2009elements, marsland2011machine, flach2012machine}.

@@

\subsection{Machine learning for security and fault management}
There are two main ways in which  machine learning enters into managing security and faults in networks. The first is \textit{using} machine learning techniques to predict and detect security breeches and faults in a network. These include machine learning algorithms for anomaly detection. The second is studying the security vulnerabilities of machine learning algorithms themselves, as the presence of adversaries becomes natural in a network setting, where real-life machine learning algorithms will be deployed. The latter is known as adversarial machine learning.

@@

When there are security breeches in classical data on a network, one desires the ability to predict and detect these disturbances, as well as a method for making one's protocols more robust against these adversaries. Machine learning is often used in anomaly detection and intrusion detection. These algorithms look for unusual data or unusual changes in data. Broadly, there are three classes of anomalies: point, contextual and collective, which refer respectively to single datum anomalies, unusual data with respect to a specified context and clusters of data which together point to an unusual pattern. Both supervised and unsupervised algorithms are used for these settings \cite{thottan2003anomaly, ahmed2007machine}. One of the prime challenges here include determining the presence of an anomaly when little data is available and determining the relevant rate of false positives and false negatives for a particular application.

Fault management in a network is also extremely relevant, especially for complex networks where there is more room for errors. One requires the prediction, detection and localisation of the fault, and most relevant machine learning methods use supervised algorithms. However, the paucity of real training data (as opposed to synthetic data generated from simulations) means that the algorithms might be poorly trained, especially in new networks \cite{hood1997proactive, kogeda2006prediction, snow2005assessing}. This is expected to be true, for instance, when a quantum internet is first set up. To accommodate for this, new methods have arisen where unsupervised machine learning techniques are used instead to detect changes in the network rather relying on labelled fault data \cite{hajji2005statistical}.

In particular, to identify and localise unusual behaviour in a network which can be due to natural faults or an adversarial party, network anomaly detection methods can be employed \cite{ahmed2007machine, fraley2017promise, joseph2013machine}. Since the results can be sensitive to the training datasets used, it is important to examine which datasets are most appropriate for one's particular applications (see a review of datasets for anomaly detection \cite{yavanoglu2017review}. In particular, there have been many proposals on using anomaly detection in the network intrusion domain. However, this approach has been criticised for its uses in real-life applications since it's often difficult to distinguish anomalies
related to intrusions from those related to other factors and the complexity of real-life networks may make it too difficult to define what is a normal signal \cite{sommer2010outside}.

@@

Machine learning algorithms themselves are vulnerable to security attacks. This comes under the field of adversarial machine learning \cite{huang2011adversarial}. There are two main types of attacks: attacks of the test data (evasion) and attacks of the training data and the machine learning models (poisoning). In real-life deployments of machine learning, data often comes from different sources which makes adversarial attacks more probable. It has been discovered that many machine learning algorithms are in fact vulnerable to adversarial attacks, the first discovered in \cite{szegedy2013intriguing}. A large proportion of the literature then focuses on the details of specific algorithms: the detection of adversaries, their different methods of attack and the particular defences to those attacks \cite{kurakin2018adversarial}. However, recently, more foundational work has emerged to try to explain the origins of this vulnerability as arising from the high dimensionality of the data involved \cite{goodfellow2014explaining, gilmer2018adversarial, mahloujifar2018curse}.

\subsection{Machine learning for traffic and routing management}
The effective operation of a network like an internet also requires automated management. This includes efficient means of traffic prediction, traffic classification, routing and congestion control. Machine learning algorithms have been developed for all these areas.

@@

Predicting network traffic is becoming increasingly important especially in diverse and complex networks. It is commonly addressed using time-series forecasting (TSF) methods. This can make use of either statistical analysis models or supervised machine learning methods \cite{bermolen2009support, chabaa2010identification, cortez2006internet}. Non-TSF methods also exist \cite{chen2016predicting, li2016inter}.

The most commonly-used technique for traffic classification is the so-called flow feature-based technique. This takes into account information on unidirectional packets sent in the network. Supervised machine learning is found to be accurate in traffic classification in this domain. However, unsupervised techniques are found to be more robust. So the joint application of both supervised and unsupervised machine learning techniques have been found to be more powerful \cite{erman2007offline, zhang2015robust}.

@@

Machine learning is most applicable to routing problems where the network is dynamical and thus requires fast updating of new optimal routes. otherwise, already-made routing tables can be sufficient for networks. Since these settings depend on frequent updating on new data, reinforcement learning algorithms are the most appropriate. In particular, Q-learning has performed well in various different networks \cite{wang2006adaptive, forster2007froms, arroyo2007q}.

@@

Congestion control in a network is also important to ensure the stability of a network and the minimisation of packet loss. Well-known congestion control methods like queue management already exist. However machine learning can be used to enhance the effectiveness of congestion control in various different networks, although mostly for TCP/IP networks \cite{liu2002end, barman2004model, el2005improving}.

%\noindent $\bullet$ Resource allocation using ML. 
% network traffic control systems \cite{fadlullah2017state}

@@

Distributed machine learning is the scenario where communication between computing devices are necessary to execute the machine learning algorithm of interest.

These distributed algorithms can be executed over an internet and become highly relevant in the following scenarios:
\begin{itemize}
\item When data used in the training and/or testing of the machine learning algorithm originates from different sources. This is the naturally-distributed setting. In these cases, it is also possible for the loading of all data onto one machine and transferring data may be too costly and interrupts the workflow. It can even be the case that this loading time is greater than computational time, so distributed computation is more resource-efficient.
\item When data is too large to be stored on the RAM of a single machine.
\item When fault tolerance becomes important. For instance, for sensitive data, if data is stored in multiple places, there is less likelihood of data corruption from any single source and the data would still be available if one source fails.
\end{itemize}

The toolbox and infrastructure for distributed machine learning is currently in rapid development. There are many algorithms available. For instance, see \cite{peteiro2013survey} for a survey. Also for distributed clustering see \cite{florian2013}. There are also systems available to cater to distributed machine learning, like MLbase \cite{MLbase}, Hadoop \cite{white2012hadoop} and Spark \cite{shanahan2015large}.

However, cautious usage is required. For example, here are some cases when one \textit{shouldn't} use distributed machine learning:
\begin{itemize}
\item When communication and synchronisation between the distributed parties provides a bottleneck for the computation.
\item When writing and running a distributed program is too complicated.
\item When one can run the same algorithm on a multi-core machine. This is possible with smart data sampling, offline schemes and efficient parallel codes. 
\end{itemize}

@@

There are at least three broad ways in which we can employ quantum resources for classical data over a network:
\begin{itemize}
\item Using quantum resources to enhance data processing at individual nodes.
\item Using quantum resources to assist in security.
\item Using quantum resources to assist in communication.
\end{itemize}

As we saw in the previous section, machine learning algorithms come in a variety of ways to assist in classical information processing tasks highly applicable in the network setting. In the presence of a classical network, our first question is then whether or not quantum resources can assist in any of the machine learning algorithms that may be relevant in a network setting . These would include tasks related to security. These belong to the class of quantum-enhanced machine learning algorithms.

In the presence of a quantum network with only classical data, a communication complexity advantage is possible \cite{brassard2003quantum}. It is unclear if and how machine learning can be helpful or relevant in this setting. However, there are already some enticing clues \cite{kane2017communication, balcan2012distributed, conitzer2004communication} on the connection between communication complexity and machine learning for classical data over a classical network. We leave questions on the roles that quantum resources can play for future investigation.

\subsection{Quantum-enhanced machine learning overview}
Quantum-enhanced machine learning algorithms are quantum algorithms performing machine learning tasks. They have so far mostly concentrated on quantum speed-ups with respect to dimensionality of the data involved. \\

@@

The first set of these algorithms have chiefly relied on assuming completely quantum devices, keeping coherence throughout the computation and could require full fault-tolerance. For those algorithms claiming up to exponential quantum speedups for supervised algorithms (see \cite{biamonte2017quantum, ciliberto2018quantum}),  the HHL algorithm \cite{harrow2009quantum} for matrix-inversion is often used. However, HHL has a number of drawbacks that make them impractical for near-term quantum devices:
\begin{itemize}
\item The ability to efficiently encode the classical data into quantum states and into quantum memory \cite{aaronson2015read}.
\item Effective read-out of the final quantum state \cite{aaronson2015read}.
\item Generally requiring high circuit-depth.
\item Restrictions on the sparsity and condition numbers of the matrix.
\end{itemize}

>>>>

\documentclass[twocolumn, aps, rmp, amsmath, amssymb, nofootinbib, superscriptaddress, longbibliography, floatfix, table-of-contents, eqsecnum]{revtex4}

@@

Today, classical machine learning is affecting the understanding and better regulation of the classical networking infrastructure underpinning the modern internet. This includes pattern recognition for individual computing devices, security and fault management, routing and traffic management, resource management, and distributed computation. In these applications, it is not only the speed of processing that is important, but reliability and security can also be paramount.

The large intersection between machine learning and network systems (the internet included) is perhaps not surprising. Firstly, machine learning relies on access to data and in many real-life applications, data naturally arises from distributed sources. Secondly, especially for complex systems like large networks, the information to process is complex, containing many uncertainties, and subject to errors. Exactly solvable models in these regimes are few and far between, and machine learning may be helpful in making predictions in these complex, messy environments.

In the coming quantum era we can envision three distinct ways where quantum resources might be introduced: quantum communication; quantum processing at individual nodes; and, data that is inherently quantum in nature. To begin, we first make a classification of the four different foreseeable types of networks, see Tab.~\ref{tab:CommTable}. These can be classified as:

\begin{itemize}
\item CC: classical data and processing over a classical network (e.g the present-day internet).
\item CQ: classical data and processing over a quantum network.
\item QC: quantum data and processing over a classical network.
\item QQ: quantum data and processing over a quantum network -- a fully-quantum internet.
\end{itemize}

Now we can pose the inevitable question: how do these different quantum resources in a network relate to machine learning?

This is not yet an active research area in its own right. However, there are some preliminary toolkits we can consider that are starting to be developed in the new field of quantum machine learning. By first summarising the intersection between the classical internet and classical machine learning, we gain insight into the kinds of tools required to start examining their quantum counterparts. For instance, we will see how aspects of quantum processing of quantum data might be aided by machine learning, how machine learning may be enhanced by quantum resources, and also in turn how machine learning may be implemented in these distributed quantum settings. See Tab.~\ref{tab:CommTable2} for a brief overview.

%It is natural to consider machine learning when we have an internet. Firstly, in machine learning applications, data naturally arise from distributed sources, so we begin from a network setting. There are also four main requirements we have in network settings (i) security (ii) speed and reliability of device information processing at each node (iii) speed and reliability of routing protocols (iv) speed and reliability of data transmission. Some of these may benefit from machine learning techniques. We then introduce quantum resources for two main reasons. The first is when the data we want to process is naturally quantum, so we wish to use machine learning to either diminish our resources to process this data or want to perform machine learning algorithms directly on this data. The second reason is that we exploit quantum resources to enhance the processing of classical data by means of machine learning.
%There are 3 parts to machine learning: data collection (offline or online); feature engineering and model learning.

\section{Classical machine learning in classical networks}

To create and maintain efficient classical networks, one requires efficient and reliable processing at individual nodes, secure processing, efficient routing and data transmission, efficient uses of resources, and a means for distributed information processing. For an overview of methods see for example \cite{boutaba2018comprehensive, wang2018machine}. We present a brief overview on how machine learning can be used in each of these areas.

\subsection{Machine learning basics for individual processors}

Machine learning algorithms allow us to make predictions about a current or future dataset without requiring explicit instructions. Since its aims are directed more towards \textit{prediction} rather than purely \textit{estimation}, it differs from the field of statistical estimation, although it shares many techniques in common.

There are three mains paradigms for machine learning:
\begin{itemize}
	\item Supervised learning: relies upon training data from which inferences and predictions about new test data can be extracted.
	\item Unsupervised learning: makes inferences from the data at hand without access to a training stage, much like a student learning without an instructor.
	\item Reinforcement learning:	operates using a different framework and aims to find the best action steps in a particular environment to maximise a given reward.
\end{itemize}

Machine learning is used regularly for data collection, feature engineering, and model learning. There are many excellent introductory texts on this topic \cite{bishop2006pattern, shalev2014understanding, trevor2009elements, marsland2011machine, flach2012machine}.

@@

\subsection{Machine learning for security and fault management}

There are two main ways in which machine learning enters into managing security and faults in networks. The first is \textit{using} machine learning techniques to predict and detect security breeches and faults in a network. These include machine learning algorithms for anomaly detection. The second is studying the security vulnerabilities of machine learning algorithms themselves, as the presence of adversaries becomes natural in a network setting, where real-life machine learning algorithms will be deployed. The latter is known as adversarial machine learning.

@@

When there are security breeches in a classical network, one desires the ability to predict and detect these disturbances, as well as a method for making one's protocols more robust against adversaries. Machine learning is often used in anomaly detection and intrusion detection. These algorithms look for unusual data or changes in the data. Broadly, there are three classes of anomalies: point, contextual, and collective, which refer respectively to single datum anomalies, unusual data with respect to a specified context, and clusters of data which together point to an unusual pattern. Both supervised and unsupervised algorithms are used in these settings \cite{thottan2003anomaly, ahmed2007machine}. One of the prime challenges here is determining the presence of an anomaly when little data is available, and determining the relevant rate of false positives and negatives for a particular application.

Fault management in a network is also extremely relevant, especially for complex networks where there is more room for errors. One requires the prediction, detection and localisation of the fault, and most relevant machine learning methods use supervised algorithms. However, the paucity of real training data (as opposed to synthetic data generated from simulations) means that the algorithms might be poorly trained, especially in newly established networks \cite{hood1997proactive, kogeda2006prediction, snow2005assessing}. This is to be reasonably anticipated with the deployment of the future quantum internet. To accommodate for this, new methods have arisen where unsupervised machine learning techniques are used instead to detect changes in the network rather than relying on labelled fault data \cite{hajji2005statistical}.

In particular, to identify and localise unusual behaviour in a network, which can be due to natural faults or an adversarial party, network anomaly detection methods can be employed \cite{ahmed2007machine, fraley2017promise, joseph2013machine}. Since the results can be sensitive to the training datasets used, it is important to examine which datasets are most appropriate for one's particular applications [see \cite{yavanoglu2017review} for a review]. In particular, there have been many proposals for utilising anomaly detection in network intrusion. However, this approach has been criticised for its use in real-world applications, where it's often difficult to distinguish anomalies
related to intrusions from those attributed to other factors, and the complexity of real-world networks may make it too difficult to define what a normal signal is \cite{sommer2010outside}.

@@

Machine learning algorithms themselves are vulnerable to security attacks. This comes under the field of adversarial machine learning \cite{huang2011adversarial}. There are two main types of attacks:
\begin{itemize}
\item Evasion: directed at the test data.
\item Poisoning: directed at the training data and machine learning models.
\end{itemize}
In real-life scenarios, data often originates from different sources, making adversarial attacks more likely. It has been discovered that many machine learning algorithms are in fact vulnerable to adversarial attacks, the first discovered in \cite{szegedy2013intriguing}. A large proportion of the literature focuses on the details of specific algorithms: the detection of adversaries; their different methods of attack; and, the particular defences against them \cite{kurakin2018adversarial}. However, recently, more foundational work has emerged, explaining the origins of this vulnerability as arising from the high dimensionality of the underlying data \cite{goodfellow2014explaining, gilmer2018adversarial, mahloujifar2018curse}.

\subsection{Machine learning for traffic and routing management}

The effective operation of a large-scale network also requires automated management. This includes efficient means for traffic prediction, traffic classification, routing, and congestion control. Machine learning algorithms have been developed for all of these.

@@

Predicting network traffic is becoming increasingly important, especially in diverse and complex networks. This is commonly addressed using time-series forecasting (TSF) methods. This can make use of either statistical analysis models or supervised machine learning methods \cite{bermolen2009support, chabaa2010identification, cortez2006internet}. Non-TSF methods also exist \cite{chen2016predicting, li2016inter}.

The most commonly-used technique for traffic classification is the so-called flow feature-based technique. This takes into account information on unidirectional packets sent across the network. For this, supervised machine learning has been found to be accurate for traffic classification. However, unsupervised techniques have been found to be more robust. Their joint application is a very powerful technique \cite{erman2007offline, zhang2015robust}.

@@

Machine learning is most applicable to dynamic routing problems, requiring rapid updating of optimal routes. Otherwise, existing routing tables can be sufficient. Since these settings depend on frequent recalculation, reinforcement learning algorithms are the most appropriate. In particular, Q-learning has performed well in various networks \cite{wang2006adaptive, forster2007froms, arroyo2007q}.

@@

Congestion control in a network is important to ensure stability and the minimisation of packet loss. Well-known congestion control methods like queue management already exist. However, machine learning can be used to enhance the effectiveness of congestion control in various scenarios, especially for TCP/IP networks \cite{liu2002end, barman2004model, el2005improving}.

@@

Distributed machine learning is simply the fusion of distributed computation and machine learning, where the learning algorithm is implemented across a network in a distributed fashion. They become highly relevant in the following scenarios:
\begin{itemize}
\item Training and/or testing data originates from different sources. This is the naturally-distributed setting.% Then, it is also possible for the loading of all data onto one machine and transferring data may be too costly and interrupts the workflow. It can even be the case that this loading time is greater than computational time, so distributed computation is more resource-efficient.
\item There is too much data to store locally on a single device.
\item When fault tolerance becomes important. For instance, for high-value data, decentralised storage provides enhanced integrity.
\end{itemize}

The toolbox and infrastructure for distributed machine learning is rapidly developing, and there are many known algorithms \cite{peteiro2013survey, florian2013}. Existing platforms catering for distributed machine learning include MLbase \cite{MLbase}, Hadoop \cite{white2012hadoop}, and Spark \cite{shanahan2015large}.

Caution is required, however, as there are cases when one \textit{shouldn't} employ distributed machine learning, such as when:
\begin{itemize}
\item Communication and synchronisation between the distributed parties provides a bottleneck for the computation.
\item Writing and running a distributed program is too complicated.
\item One can run the same algorithm on a multi-core machine. This is possible with smart data sampling, offline schemes, and efficient parallel codes.
\end{itemize}

@@

There are at least three broad ways in which we can employ quantum resources for classical data over a network:
\begin{itemize}
\item Quantum resources enhance data-processing at individual nodes.
\item Quantum resources improve security.
\item Quantum resources enhance communication.
\end{itemize}

In a classical network, the first question is whether or not quantum resources can assist in any of the relevant machine learning algorithms, including improving security. These belong to the class of quantum-enhanced machine learning algorithms.

In a quantum network with only classical data, communication complexity improvements are possible \cite{brassard2003quantum}. It is unclear whether machine learning has utility in this setting, although there are some promising clues \cite{kane2017communication, balcan2012distributed, conitzer2004communication} on the connection between communication complexity and machine learning.

\subsection{Quantum-enhanced machine learning overview}

Quantum-enhanced machine learning algorithms are quantum algorithms performing machine learning tasks. They have so far mostly concentrated on quantum speed-ups with respect to the dimensionality of the underlying data.

@@

The first of these algorithms chiefly relied on assuming completely quantum devices, maintaining coherence throughout the computation, and could require full fault-tolerance. For algorithms claiming exponential quantum ehnhancement for supervised learning \cite{biamonte2017quantum, ciliberto2018quantum}, the HHL algorithm \cite{harrow2009quantum} for matrix inversion is often employed. However, HHL has a number of drawbacks, making it impractical for near-term quantum devices:
\begin{itemize}
\item The ability to efficiently encode classical data into quantum states and quantum memory \cite{aaronson2015read}.
\item Effective quantum state read-out \cite{aaronson2015read}.
\item They generally require high circuit-depth.
\item There are restrictions on the sparsity and conditioning of matrices.
\end{itemize}

% Peter up to here
